---
title: "The Attractive Flu Shot 2019 - Analysis"
author: "Amnon Maltz and Adi Sarid"
date: "9/13/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Background

The following report summarizes the analysis of an "attractive flu shot" experiment. This is merely a technical report including all the data import, transformations, modelling, visualization. 

The report is built in the spirit of "reproduceable research", i.e., assuming a similar data format, all the statistical procedures can be performed in standard packages available in R.

```{r data import}

library(tidyverse)

# To use the saridr package use:
# devtools::install_github("sarid-ins/saridr")

Sys.setlocale("LC_ALL", "Hebrew") # some responses are in Hebrew - making sure encoding is read correctly.

flu_data <- read_csv("data/flu_data_final.csv") %>% 
  mutate(treatment = fct_relevel(treatment, "Control", "Recommendation", "Stock", "Cost", "Benefit"))

```

## Validation check

We check for the following inconsistencies:

   * Respondents which checked the "vaccinated last year", but have less than 1 vaccinations in the last 5 years.
   * Respondents which did not check the "vaccinated last year", but have 5 vaccinations in the last 5 years.

```{r validation check}

flu_data %>% 
  filter(last_year_vaccine == 1 & five_years_vaccinations == 0 | 
           last_year_vaccination == 0 & five_years_vaccinations == 5)

```

There are only 3 such inconsistencies, after reading their open-ended action explanation, they all seem to be answering seriously, so we decided to keep them.

## Examing the level of certainty depending on treatment

We show the distribution of certainty level as a function of the treatment presented to the respondents.


```{r certainty versus treatment}

ggplot(flu_data, aes(certainty)) + 
  geom_histogram(bins = 5) + 
  facet_wrap(~treatment)

```

## Examine correlation of number of vaccinations versus certainty

We examine the relationship between the number of vaccinations in the last 5 years and the certainty level. Logic dictates that respondents which vaccinated 5 times or 0 times are much more "certain" than those who vaccinated 1-4 times.



```{r vaccination history and certainty}
history_vs_certainty <- flu_data %>% 
  mutate(vaccination_history = case_when(five_years_vaccinations %in% c(5) ~ 3,
                                         five_years_vaccinations %in% c(0:2) ~ 1,
                                         five_years_vaccinations %in% c(3:4) ~ 2)) %>% 
  mutate(certainty_3_levels = case_when(certainty >=4 ~ 3,
                                        certainty == 3 ~ 2,
                                        certainty <= 2 ~ 1))

cor(history_vs_certainty %>%
      select(five_years_vaccinations, certainty, certainty_3_levels, vaccination_history)) %>% 
  knitr::kable()
```


## Effect of treatment split by certain/uncertain (Table 3 in the paper) 

The table compares the effect of the treatment by certainty level.

```{r treatment by certainty}
total_vaccinated <- flu_data %>% 
  group_by(treatment) %>% 
  summarize(tot_vaccinated = mean(action_binary)) %>% 
  spread(treatment, tot_vaccinated) %>% 
  mutate(certainty = NA)

# table in percentages

flu_data %>% 
  group_by(treatment, certainty) %>% 
  summarize(prop_vaccinated = mean(action_binary)) %>% 
  spread(treatment, prop_vaccinated) %>% 
  bind_rows(total_vaccinated) %>% 
  knitr::kable()

# Table in absolute numbers

flu_data %>% 
  group_by(treatment, certainty) %>% 
  summarize(prop_vaccinated = mean(action_binary),
            vaccinated_abs = sum(action_binary),
            total_cell = n()) %>% 
  mutate(tbl_val = paste0(round(prop_vaccinated*100), "% (", vaccinated_abs, "/", total_cell, ")")) %>% 
  select(treatment, certainty, tbl_val) %>% 
  spread(treatment, tbl_val) %>% 
  knitr::kable()
  
# Splitting the certainty to two levels (<=3 versus >=4)
flu_data %>% 
  mutate(certainty_binary = certainty <= 3) %>% 
  group_by(treatment, certainty_binary) %>% 
  summarize(prop_vaccinated = mean(action_binary),
            vaccinated_abs = sum(action_binary),
            total_cell = n()) %>% 
  mutate(tbl_val = paste0(round(prop_vaccinated*100), "% (", vaccinated_abs, "/", total_cell, ")")) %>% 
  select(treatment, certainty_binary, tbl_val) %>% 
  spread(treatment, tbl_val) %>% 
  mutate(certainty_binary = ifelse(certainty_binary, "Uncertain", "Certain")) %>% 
  knitr::kable()


```

## Checking for two-way ANOVA

We examine the two-way anove with vaccination action as the dependent variable versus treatment and certainty. We are using a dichotomous variable certain/uncertain (where uncertain = certainty<=3).

```{r two way anova}
aov_model <- aov(data = flu_data, formula = action_binary ~ (certainty<=3)*treatment) 
summary(aov_model)
```

## One way ANOVA

```{r one way anova}
aov_uncertain <- aov(data = flu_data %>% filter(certainty <= 3), 
                     formula = action_binary ~ treatment) 

aov_certain <- aov(data = flu_data %>% filter(certainty >= 4), 
                     formula = action_binary ~ treatment) 
summary(aov_uncertain)
summary(aov_certain)
```


## Multiple comparisons using Dunnettes test

```{r mult comp Dunnettes test}

dunnette <- multcomp::glht(aov_uncertain,
                          linfct = multcomp::mcp(treatment = 
                                                   c("Recommendation - Control = 0",
                                                     "Stock - Control = 0",
                                                     "Cost - Control = 0",
                                                     "Benefit - Control = 0"))) %>% 
  summary()

dunnette

```

## Effect of treatment split by number of vaccinations in the last 5 years

```{r treatment by vaccination history}
total_vaccinated <- flu_data %>% 
  group_by(treatment) %>% 
  summarize(tot_vaccinated = mean(action_binary)) %>% 
  spread(treatment, tot_vaccinated) %>% 
  mutate(five_years_vaccinations = NA)

flu_data %>% 
  group_by(treatment, five_years_vaccinations) %>% 
  summarize(prop_vaccinated = mean(action_binary)) %>% 
  spread(treatment, prop_vaccinated) %>% 
  bind_rows(total_vaccinated) %>% 
  knitr::kable()

flu_data %>% 
  group_by(treatment, five_years_vaccinations) %>% 
  summarize(tot_vaccinated = mean(action_binary)) %>% 
  ungroup() %>%
  ggplot(aes(x = five_years_vaccinations, y = tot_vaccinated, fill = treatment)) + 
  geom_col(position = "dodge", color = "black") + 
  theme_bw() + 
  scale_fill_brewer(palette = "Dark2") + 
  xlab("Number of vaccinations in the last 5 years") + 
  ylab("% respondents willing to vaccinate next year")
  

```


## Effects of treatment as a function of income

We examine the influence of income on the effect of the treatments

```{r income treatment interaction}

flu_data %>% 
  group_by(treatment, income) %>% 
  summarize(tot_vaccinated = mean(action_binary)) %>% 
  spread(treatment, tot_vaccinated) %>% 
  slice(c(2, 4, 3, 5, 6, 1)) %>% 
  knitr::kable()

flu_data %>% 
  group_by(certainty) %>% 
  saridr::prop(income) %>% 
  spread(certainty, prop) %>% 
  knitr::kable()


```

## Demographics (Table 1)

The demographics of participants

```{r demographics}
gender <- flu_data %>% 
  group_by(treatment) %>% 
  saridr::prop(gender) %>% 
  spread(treatment, prop) %>% 
  filter(gender == "Female") %>% 
  mutate(Demography = paste0("% ", gender)) %>% 
  select(-gender) %>% 
  mutate_if(is.numeric, ~{paste0(round(.*100,1), "%")})

age <- flu_data %>% 
  group_by(treatment) %>% 
  summarize(mean = mean(age, na.rm = T),
            sd = sd(age, na.rm = T)) %>% 
  mutate(tbl_value = paste0(round(mean, 1), " (", round(sd, 1), ")")) %>% 
  ungroup() %>%
  select(treatment, tbl_value) %>% 
  spread(treatment, tbl_value) %>% 
  mutate(Demography = "Age mean (sd)")

income <- flu_data %>% 
  group_by(treatment) %>% 
  saridr::prop(income) %>% 
  spread(treatment, prop) %>% 
  slice(c(2, 4, 3, 5, 6, 1)) %>% 
  mutate_if(is.numeric, ~{paste0(round(.*100,1), "%")}) %>% 
  rename(Demography = income)
  

education <- flu_data %>% 
  group_by(treatment) %>% 
  saridr::prop(education) %>% 
  spread(treatment, prop) %>% 
  slice(c(1:2, 4:5, 3, 6)) %>% 
  mutate_if(is.numeric, ~{paste0(round(.*100,1), "%")}) %>% 
  rename(Demography = education) %>% 
  mutate(Demography = fct_explicit_na(Demography))

demographics_latex_table <- bind_rows(gender,
                                      age,
                                      income,
                                      education) %>% 
  select(Demography, everything())

# Code to generate a nice LaTex (booktabs style) version

# knitr::kable(demographics_latex_table, format = "latex", booktabs = TRUE)

```

