---
title: "The Attractive Flu Shot 2020 - Summary Report"
author: "Amnon Maltz and Adi Sarid"
date: "2020-02-24"
output:
  html_document:
    df_print: paged
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Background

The following report summarizes the analysis of an "attractive flu shot" experiment. This is merely a technical report including all the data import, transformations, modelling, visualization. 

The report is built in the spirit of "reproduceable research", i.e., assuming a similar data format, all the statistical procedures can be performed in standard packages available in R.

```{r data import}

suppressMessages(suppressWarnings(library(tidyverse)))

# To use the saridr package use:
# devtools::install_github("sarid-ins/saridr")

# some responses are in Hebrew - making sure encoding is read correctly.
suppressMessages(Sys.setlocale("LC_ALL", "Hebrew")) 


flu_data <- read_csv("data/flu_data_final.csv") %>% 
  mutate(treatment = fct_relevel(treatment, "Control", "Recommendation", "Stock", "Cost", "Benefit")) %>% 
  mutate(age_group = cut(age, breaks = c(0, 17.9, 29, 39, 49, 59, 100)))

```

## Demographics (Table 1)

The demographics of participants

```{r demographics}
gender <- flu_data %>% 
  group_by(treatment) %>% 
  saridr::prop(gender) %>% 
  spread(treatment, prop) %>% 
  filter(gender == "Female") %>% 
  mutate(Demography = paste0("% ", gender)) %>% 
  select(-gender) %>% 
  mutate_if(is.numeric, ~{paste0(round(.*100,1), "%")})

age <- flu_data %>% 
  group_by(treatment) %>% 
  summarize(mean = mean(age, na.rm = T),
            sd = sd(age, na.rm = T)) %>% 
  mutate(tbl_value = paste0(round(mean, 1), " (", round(sd, 1), ")")) %>% 
  ungroup() %>%
  select(treatment, tbl_value) %>% 
  spread(treatment, tbl_value) %>% 
  mutate(Demography = "Age mean (sd)")

income <- flu_data %>% 
  group_by(treatment) %>% 
  saridr::prop(income) %>% 
  spread(treatment, prop) %>% 
  slice(c(2, 4, 3, 5, 6, 1)) %>% 
  mutate_if(is.numeric, ~{paste0(round(.*100,1), "%")}) %>% 
  rename(Demography = income)
  

education <- flu_data %>% 
  group_by(treatment) %>% 
  saridr::prop(education) %>% 
  spread(treatment, prop) %>% 
  slice(c(1:2, 4:5, 3, 6)) %>% 
  mutate_if(is.numeric, ~{paste0(round(.*100,1), "%")}) %>% 
  rename(Demography = education) %>% 
  mutate(Demography = fct_explicit_na(Demography))

demographics_latex_table <- bind_rows(gender,
                                      age,
                                      income,
                                      education) %>% 
  select(Demography, everything())

# Code to generate a nice LaTex (booktabs style) version

knitr::kable(demographics_latex_table, format = "markdown", booktabs = TRUE) # <- for latex change argument to: format = "latex

```

## Effect of treatment split by number of vaccinations in the last 5 years (Table 2)

```{r treatment by vaccination history}
total_vaccinated <- flu_data %>% 
  group_by(treatment) %>% 
  summarize(tot_vaccinated = mean(action_binary),
            sample_size = n()) %>% 
  mutate(prop_vaccinated_str = paste0(round(tot_vaccinated*100), "% (",
                                      tot_vaccinated*sample_size,"/", sample_size,")")) %>% 
  select(treatment, prop_vaccinated_str) %>% 
  spread(treatment, prop_vaccinated_str) %>% 
  mutate(five_years_vaccinations = "Overall") 

flu_data %>% 
  group_by(treatment, five_years_vaccinations) %>% 
  summarize(prop_vaccinated = mean(action_binary),
            sample_size = n()) %>% 
  mutate(prop_vaccinated_str = paste0(round(prop_vaccinated*100), "% (",
                                      prop_vaccinated*sample_size,"/", sample_size,")")) %>% 
  select(-prop_vaccinated, -sample_size) %>% 
  spread(treatment, prop_vaccinated_str) %>% 
  mutate(five_years_vaccinations = as.character(five_years_vaccinations)) %>% 
  bind_rows(total_vaccinated) %>% 
  knitr::kable(format = "latex")
```

## Using Chi square test to examine the independence of the treatments effectiveness per number of vaccinations over the last 5 years

```{r pearson chi sq}

chisq3 <- flu_data %>% 
  filter(five_years_vaccinations == 0) %>% 
  count(treatment, action_binary) %>% 
  pivot_wider(id_cols = treatment, names_from = action_binary, values_from = n, values_fill = list(n=0)) %>% 
  column_to_rownames(var = "treatment") %>% 
  as.matrix() %>% 
  chisq.test()

chisq4 <- flu_data %>% 
  filter(five_years_vaccinations == 1) %>% 
  count(treatment, action_binary) %>% 
  pivot_wider(id_cols = treatment, names_from = action_binary, values_from = n, values_fill = list(n=0)) %>% 
  column_to_rownames(var = "treatment") %>% 
  as.matrix() %>% 
  chisq.test()

chisq5 <- flu_data %>% 
  filter(five_years_vaccinations == 2) %>% 
  count(treatment, action_binary) %>% 
  pivot_wider(id_cols = treatment, names_from = action_binary, values_from = n, values_fill = list(n=0)) %>% 
  column_to_rownames(var = "treatment") %>% 
  as.matrix() %>% 
  chisq.test()

fisher_exact1 <- flu_data %>% 
  filter(five_years_vaccinations >= 3) %>% 
  count(treatment, action_binary) %>% 
  pivot_wider(id_cols = treatment, names_from = action_binary, values_from = n, values_fill = list(n=0)) %>% 
  column_to_rownames(var = "treatment") %>% 
  as.matrix() %>% 
  fisher.test()

chisq9 <- flu_data %>% 
  count(treatment, action_binary) %>% 
  pivot_wider(id_cols = treatment, names_from = action_binary, values_from = n, values_fill = list(n=0)) %>% 
  column_to_rownames(var = "treatment") %>% 
  as.matrix() %>% 
  chisq.test()

```

## Logistic regression of intentions to vaccinate (Table 3)

A logistic regression model of intentions to receive the vaccination.

```{r logistic regression model nominal}

flu_data_for_glm <- flu_data %>% 
  # mutate(treatment = fct_relevel(treatment, "Control", after = 4)) %>% 
  mutate(certain = certainty >= 4)

vaccination_glm0 <- glm(data = flu_data_for_glm, family = binomial, 
                        formula = action_binary ~ treatment)

summary(vaccination_glm0)
```

The same model, with interaction of traetment*certainty

```{r logistic regression with uncertain interaction}
vaccination_glm1 <- glm(data = flu_data_for_glm, 
                        family = binomial, 
                        formula = action_binary ~ treatment + five_years_vaccinations + treatment*five_years_vaccinations)

summary(vaccination_glm1)
```

A complete model including background demographic variables and vaccinated last.

```{r logistic regression with five year vaccination and demography}

vaccination_glm2 <- glm(data = flu_data_for_glm, family = binomial, 
                        formula = action_binary ~ 
                          treatment + five_years_vaccinations + treatment*five_years_vaccinations + 
                          gender + age + income + education)

summary(vaccination_glm2)

```

```{r coefficient comparison with all controls}

# recommendation - cost
rec_min_cost <- vaccination_glm2$coefficients[2] - vaccination_glm2$coefficients[4]

rec_min_cost
# var(recommendation - cost) = var(recommendation) + var(cost) - 2*cov(recommendation, cost)

vcov_glm2 <- vcov(vaccination_glm2)
rec_min_cost_std_err <- sqrt(
  vcov_glm2[2, 2] + vcov_glm2[4, 4] - 2*vcov_glm2[2, 4]
)

rec_min_cost_std_err

# compute the significance (p-value)

z_diff <- (rec_min_cost / rec_min_cost_std_err)
p_diff <- 2*(1-pnorm(abs(z_diff), 0, 1))

p_diff

```

The code to generate a stargazer latex table (Table 3).

```{r regression output}
## Disable/enable according to need.
stargazer::stargazer(vaccination_glm0,
                     vaccination_glm1,
                     vaccination_glm2,
                     type = "latex")

```

In addition to the previous test, we also compare the difference between proportions directly.

```{r recommendation versus cost vaccination proportions test}

trials <- flu_data %>% 
  filter(treatment %in% c("Recommendation",
                          "Cost")) %>% 
  count(treatment, action_binary) %>% 
  pivot_wider(id_cols = treatment, names_from = action_binary, values_from = n) %>% 
  mutate(total = `0`+`1`) %>% 
  rename(successes = `1`) %>% 
  select(-`0`)

prop.test(x = trials$successes, n = trials$total, correct = T)

```

## Open ended question - content analysis

In the menuscript, results section, the following appears in the body (not as a table).

```{r comparison comments vs treatment}
open_ended <- read_csv("data/open_ended_content_analysis.csv", skip = 1,
                       col_names = c("id", "comparison", "spam")) %>% 
  mutate_all(~ifelse(is.na(.), 0, .))

flu_data %>% 
  left_join(open_ended) %>% 
  group_by(treatment) %>% 
  summarize(portion_comparison = mean(comparison))

```

## Multinomial regression of vaccination intentions versus treatments

We treat the action states: no vaccination, late vaccination, and early vaccinations in a multinomial logistic regression framework.

```{r multinomial regression intentions}

flu_multi_action <- flu_data %>%
  filter(treatment != "Control") %>% 
  mutate(action_recoded = factor(action_recoded,
                                 levels = c("not_vaccinate", "vaccinate_late", "vaccinate_early"))) %>% 
  mutate(treatment = droplevels(treatment))

multinom_res <- nnet::multinom(formula = action_recoded ~ treatment,
               data = flu_multi_action) %>% 
  summary()

# significance test

z_multinom <- multinom_res$coefficients/multinom_res$standard.errors
p_multinom <- (1 - pnorm(abs(z), 0, 1))*2

#Nice table with coefficients per model and std. err and p-values.

multinom_tbl <- tibble(
  variable = colnames(multinom_res$coefficients),
  late_coeff = multinom_res$coefficients[1,],
  late_stderr = multinom_res$standard.errors[1,],
  late_z = z_multinom[1,],
  late_p_val = p_multinom[1,],
  early_coeff = multinom_res$coefficients[2,],
  early_stderr = multinom_res$standard.errors[2,],
  early_z = z_multinom[2,],
  early_p_val = p_multinom[2,]
)

#knitr::kable(multinom_tbl, format = "latex", booktabs = T)

```


## FDR correction for significance tests within the document

For the chi square tests (in the body and appendix).

```{r fdr correction}
p.adjust(p = c(0.02569, 0.04222, 0.4342,
               0.444966166, 0.009687003, 0.003422310, 0.224901437, 0.200187974,
               0.00796), 
         method = "fdr")
```

# Analysis provided in the supplementary material

## Supplementary material A: additional crosstabulations in higher resolution.

### Certainty level distribution

```{r certainty level}

flu_data %>% 
  saridr::prop(certainty, leave_n = T)

```


### Effect of treatment by certainty level (1-5)

```{r effect treatement by certainty levels}
flu_data %>%
  group_by(treatment, certainty) %>%
  summarize(prop_vaccinated = mean(action_binary)) %>%
  spread(treatment, prop_vaccinated) %>%
  # bind_rows(total_vaccinated %>% select(-five_years_vaccinations)) %>%
  knitr::kable()
```

### Using Chi square test to examine the independence of the intervention vs. intentions to vaccinate, per certainty level

```{r pearson chi sq per certainty}

get_chisq_pval <- function(certainty_level){
  tmp <- flu_data %>% 
    filter(certainty %in% certainty_level) %>% 
    count(treatment, action_binary) %>%
    pivot_wider(id_cols = treatment, names_from = action_binary, values_from = n, values_fill = list(n=0)) %>% 
    column_to_rownames(var = "treatment") %>% 
    as.matrix() %>% 
    chisq.test()
  
  tibble(p.val = tmp$p.value, chi_sq = tmp$statistic)
}

certainty_chisq <- list(1,2,3,4,5) %>% 
  map_df(.f = get_chisq_pval) %>% 
  mutate(certainty_level = as.character(list(1,2,3,4,5)))

p_value_test2_for_fdr <- certainty_chisq$p.val
```

### Regression with certainty as a continuous variable

```{r regression with cerainty levels}

vaccination_glm_certain_lvls1 <- glm(data = flu_data_for_glm %>% 
                                       mutate(certainty = certainty - 1), 
                                     family = binomial,
                                    formula = action_binary ~ treatment + certainty + treatment*certainty)

# summary(vaccination_glm_certain_lvls1)

vaccination_glm_certain_lvls2 <- glm(data = flu_data_for_glm %>% 
                                       mutate(certainty = certainty - 1), 
                                     family = binomial, 
                                    formula = action_binary ~ treatment + certainty + treatment*(certainty) +
                                      last_year_vaccination + gender + age + income + education)

# summary(vaccination_glm_certain_lvls2)

# stargazer::stargazer(vaccination_glm_certain_lvls1, vaccination_glm_certain_lvls2)

```

## Ordinal regression of certainty levels

The following is a request by the associate editor.

```{r ordinal regression certainty}

certainty_ordinal_regression <- 
  MASS::polr(formula = factor(certainty) ~ treatment,
             data = flu_data, method = "logistic")

summary(certainty_ordinal_regression)

# a latex table with the coefficients
ord_log <- summary(certainty_ordinal_regression)$coefficients %>% 
  as.data.frame() %>% 
  rownames_to_column() %>% 
  rename(variable = rowname) %>% 
  mutate(p_value = 2*(1-pt(abs(`t value`), df = NROW(flu_data) - 8)))

# knitr::kable(ord_log, format = "latex", booktabs = T, digits = 2)
```

The ordinal logistic equation assumes the following model:

$$\zeta_0=-\infty<\zeta_1<\ldots<\zeta_5<\zeta_6=\infty$$

With:

$$\operatorname{logit}P(Y\leq k|x)=\zeta_k-\eta x$$


## Supplementary material B: treatment effect by income level

Is there an effect of income on the increased costs or the benefit treatments?

```{r income effect}

flu_attraction_income <- flu_data %>% 
  # filter(treatment %in% c("Control", "Cost", "Benefit")) %>% 
  select(treatment, action_binary, income) %>% 
  filter(income != "(Missing)") %>% 
  mutate(income = recode_factor(income, 
                                "Under" = "Under",
                                "Slightly under" = "Under",
                                "Average" = "Average",
                                "Slightly above" = "Above",
                                "Above" = "Above"))

flu_attraction_income %>% 
  group_by(treatment, income) %>% 
  summarize(vaccination_intentions = mean(action_binary)) %>% 
  pivot_wider(names_from = treatment, values_from = vaccination_intentions) %>% 
  rename(Income = income) %>% 
  mutate_if(is.numeric, ~paste0(round(.*100), "%"))

# overall chisquare per each income level considering all treatments

chi_square_goodness_of_fit_income <- function(income_lvl){
  tmp <- flu_attraction_income %>% 
    filter(income %in% income_lvl) %>%  
    count(treatment, action_binary) %>% 
    pivot_wider(names_from = action_binary, values_from = n) %>% 
    select(2:3) %>% 
    as.matrix() %>% 
    chisq.test()
  
  tibble(p_val = tmp$p.value, chi = tmp$statistic)
  
}

chi_goodness_income <- tibble(income_lvl = c("Under", "Average", "Above")) %>% 
  mutate(chi_sq = map(income_lvl, chi_square_goodness_of_fit_income)) %>% 
  unnest(chi_sq)

chi_goodness_income

# p_value_test3_for_fdr <- chi_goodness_income$p_val

#knitr::kable(chi_goodness_income, format = "markdown")

```

## Supplementary material C: Data Quality Assessment

We check for the following inconsistencies:

   * Respondents which checked the "vaccinated last year", but have less than 1 vaccinations in the last 5 years.
   * Respondents which did not check the "vaccinated last year", but have 5 vaccinations in the last 5 years.

```{r validation check}

flu_data %>% 
  filter(last_year_vaccination == 1 & five_years_vaccinations == 0 | 
           last_year_vaccination == 0 & five_years_vaccinations == 5) %>% 
  select(id, started, submitted, status, treatment, open_ended_action_explanation,
         certainty, last_year_vaccination, five_years_vaccinations)

```

There are only 12 such inconsistencies, after reading their open-ended action explanation, they all seem to be answering seriously, so we decided to keep them.

## Open ended content analysis - detection of spam

The open ended questions were read and a minor portion (10 out of 3271) were classified as spam.

```{r open ended}

open_ended %>% 
  count(spam)

```

## Supplementary material E: Distribution of the sample by age

```{r distribution sample age}

flu_data %>% saridr::prop(age_group, leave_n = T)
  
```

# Additional materials (only mentioned in the paper)

## Time to complete questionnaire

```{r time to complete}
complete_duration <- flu_data %>% 
  mutate_at(vars(started, submitted), ~lubridate::dmy_hms(.)) %>% 
  mutate(duration = submitted - started) %>% 
  select(id, duration) %>% 
  mutate(duration_groups = cut(as.numeric(duration), c(0, 30, 60, 120, 500, 10000)))

ggplot(complete_duration, aes(duration)) + 
  stat_ecdf() + 
  theme_bw() + 
  coord_cartesian(xlim = c(0, 500))

quantile(complete_duration$duration, c(0.05, 0.95))

ggplot(complete_duration, aes(duration)) + 
  geom_histogram(bins = 30) + 
  scale_x_continuous(limits = c(0, 500))

complete_duration %>% 
  count(duration_groups) %>% 
  mutate(prop = n/sum(n))

```

## Number of late vaccinations per treatment

This examines the number of late vaccinations in each treatment group.

```{r late vaccinations}

flu_data %>% 
  group_by(treatment) %>% 
  saridr::prop(action_recoded) %>% 
  pivot_wider(id_cols = treatment, names_from = action_recoded, values_from = c(prop)) %>% 
  knitr::kable("markdown")

flu_data %>% 
  group_by(treatment) %>% 
  count(action_recoded) %>% 
  pivot_wider(id_cols = treatment, names_from = action_recoded, values_from = n) %>% 
  knitr::kable("markdown")


```

## Percent of vaccinated last year crosstabulated with certainty level

```{r vaccinated last year vs certainty}

flu_data %>% 
  group_by(certainty) %>% 
  summarize(percent_vaccinated_last_year = mean(last_year_vaccination, na.rm = T)) %>% 
  knitr::kable(format = "markdown")

cor(flu_data$certainty, flu_data$last_year_vaccination, use = "pairwise.complete.obs")
cor(flu_data$certainty, flu_data$five_years_vaccinations, use = "pairwise.complete.obs")
cor.test(flu_data$certainty, flu_data$five_years_vaccinations, 
         use = "pairwise.complete.obs")

flu_data %>% 
  group_by(certainty) %>% 
  saridr::prop(certainty, five_years_vaccinations, leave_n = T) %>% 
  rename(n = n4prop) %>% 
  pivot_wider(names_from = five_years_vaccinations, values_from = c(n, prop)) %>% 
  knitr::kable(format = "markdown")


flu_data %>% 
  ggplot(aes(x = factor(certainty), fill = factor(five_years_vaccinations))) + 
  geom_bar(position = "fill", color = "black") + 
  theme_bw() + 
  scale_fill_brewer() + 
  scale_y_continuous(labels = scales::percent) + 
  ylab("Population proportion") + 
  xlab("Certainty level") + 
  guides(fill = guide_legend("Vaccination in\npast five years"))

```

## Addition - correlation coefficient of vaccination history versus certainty

```{r vaccination history vs certainty}
cor(flu_data$five_years_vaccinations, flu)
```

## Late vaccinations

Those who chose to vaccinate later rather than earlier - proportions overall and in each treatment group.

```{r late vaccinators}

flu_data %>% 
  filter(treatment != "Control") %>% 
  saridr::prop(action_recoded, leave_n = T)

flu_data %>% 
  group_by(treatment) %>% 
  saridr::prop(action_recoded, leave_n = T)
  
```

## Vaccination history: last year versus five years

The correlation between five year vaccination history, and last year's vaccination

```{r vaccination history correlation}
cor(flu_data$five_years_vaccinations, flu_data$last_year_vaccination, use = "pairwise.complete.obs")
ggplot(flu_data, aes(x = factor(five_years_vaccinations), fill = factor(last_year_vaccination))) + 
  geom_bar(position = "fill") + 
  theme_bw()
```



## Examining the level of certainty level as a function of treatment type

We show the distribution of certainty level as a function of the treatment presented to the respondents. This chart did not make it into the manuscript.

```{r certainty versus treatment}

ggplot(flu_data, aes(certainty)) + 
  geom_histogram(bins = 5) + 
  facet_wrap(~treatment)

```

We want to make sure that the treatment type did not affect the reported certainty level, via $\chi^2$ test.

```{r certainty distribution versus treatment}
flu_data %>% 
  count(treatment, certainty) %>% 
  pivot_wider(id_cols = treatment, names_from = certainty, values_from = n, values_fill = list(n=0)) %>% 
  column_to_rownames(var = "treatment") %>% 
  as.matrix() %>% 
  chisq.test()
```
